---
title: "R Cheatsheet"
output: html_document
---
Kate's personal log of interesting and useful R snippets

Last updated: 23Feb16

## Data Wrangling
### Data Input Basics
Creating a simple dataframe for example purposes
```{r}
Location <- c("New York", "London", "new york", "London", "London")
Product <- c("Tea", "Coffee", "Coffee", NA, "Coffee")
Revenue <- c("USD 10000", "5000", "6,500", "$7850", "300")
YearFounded <- c(1985, 2001, NA, NA, 1996)
SampleSet <- data.frame(Location, Product, Revenue, YearFounded, stringsAsFactors=FALSE)
SampleSet
```

Saving a copy of the dataframe as .csv (comma-separated) data
```{r}
write.csv(SampleSet, "SampleSet.csv")
```

Reading in .csv (comma-separated) data
```{r}
NewSampleSet <- read.csv("SampleSet.csv")

NewSampleSet
```

Reading in .tsv (tab-separated) data
```{r}
Football <- read.table("Football.tsv", header=TRUE, sep = "\t")
```

### Looking at Basic Structure
Show the dimensions of the dataframe
```{r}
dim(SampleSet)
```

Show the top few rows of content in a dataframe
```{r}
head(Football)
```

Show the class of each field in a dataframe
```{r}
sapply(SampleSet, class)
```

Show a simple summary of each field in a dataframe
```{r}
summary(Football)
```

See the distributions of values for a certain field
```{r}
table(SampleSet$Location)
```

List all unique values for a certain field
```{r}
unique(SampleSet$Location)
```

### Column Headings
Listing all column headings
```{r}
names(SampleSet)
```

Substituting new column names using rename()
```{r}
#Creating a vector of new column names (in any order), with old column names as the labels
FootballNames <- c("Y"= "Points",
           "X5" = "Shots",
           "X7" = "Box",
           "X11" = "Passes",
           "X15" = "Crosses",
           "X" = "Team",
           "X18" = "GoalsConceded",
           "X23" = "YellowCards")
FootballNames
#Load the 'plyr' package
library(plyr)
#Replace existing column names using plyr package
Football <- rename(Football, FootballNames)
#Show top lines of data to verify it has been read correctly
head(Football)
```

### Understanding structure of individual columns
See now many non-NA values are contained within a column
```{r}
length(which(!is.na(SampleSet$Product)))
```

### Cleaning up content
Duplicating a field in the dataset (but giving it a new name)
```{r}
SampleSet$OtherLocation <- SampleSet$Location
SampleSet
```

Delete a field from the dataset
```{r}
SampleSet$OtherLocation <- NULL
SampleSet
```

Convert all contents of a column to proper case
```{r}
SampleSet$Location

#Add a function to make proper case
properCase <- function(x) {
  lower <- tolower(x)
  s <- strsplit(lower, " ")[[1]]
  paste(toupper(substring(s, 1,1)), substring(s, 2),
      sep="", collapse=" ")
}
#Apply this function to a column
SampleSet$LocationProper <- sapply(SampleSet$Location, properCase)

SampleSet$LocationProper
```

Extract the digits in a column and make column into an integer column (by replacing all non-digit characters with blank space)
```{r}
SampleSet$Revenue <- as.integer(gsub("[^0-9]", "", SampleSet$Revenue))
SampleSet
```

## Grouping and Summarising
### Using dplyr
```{r}
library(dplyr)
summarise(group_by(SampleSet, LocationProper), count = n(), MeanRev = mean(Revenue, na.rm=TRUE))
```


## Plots
### Simple Plots in Base R
Show a histogram of the distribution of values
```{r}
hist(SampleSet$Revenue)
```

Same as previous, but with a lot more bins to the histogram
```{r}
hist(SampleSet$Revenue, breaks=100)
```

Boxplot of a variable
```{r}
boxplot(Football$Points, ylab = "Points Scored")
```

Boxplot of a variable, separated by the factor categories of another variable
```{r}
boxplot(Football$Points ~ Football$Tiers, ylab = "Points Scored", xlab = "Tier")
```

Plot a QQ plot to see if a field is distributed normally
```{r}
qqnorm(Football$Shots, main = "Normal Q-Q Plot: Ratio of Short to Long Passes")
qqline(Football$Shots)
```

### Plots with ggplot2

### Fancier Plots from Other Packages
A pretty scatterplot matrix using the *cars* package
```{r}
library(car)
spm(~ Points + Shots + Box + Passes, data = Football, smoother = FALSE)
```

Same as above but where factors from another fields are used to colour-code the dots, and with smoothed trend lines
```{r}
library(car)
spm(~ Points + Shots + Box + Passes | Tiers, data = Football)
```

## Statistical Tools
### Statistical Tests and ANOVA
Shapiro-Wilk test for normal distribution
```{r}
shapiro.test(Football$Passes)
```

ANOVA One-Way test on a variable across categories (described by a factor fariable)
```{r}
oneway.test(Football$Points ~ Football$Tiers)
```

### Correlation
Output a Pearson correlation matrix (after reducing the default number of digits shown)
```{r}
options(digits = 3)
cor(Football[,2:5])
#Reset digits options back to original
options(digits = 7)
```

Plot a visual form of the Pearson correlation matrix using the *ellipse* package
```{r}
library(ellipse)
CorMatrix <- cor(Football[,2:5])
#Make the plot but modify the margins slightly
plotcorr(CorMatrix, mar = c(0.1, 0.1, 0.5, 0.1))
```

Test the correlation between two fields (Null hypothesis: no correlation)
```{r}
cor.test(Football$Points, Football$Passes)
```

### Linear Regression
Building a linear regression model and summarising it
```{r}
LinearReg <- lm(Points ~ Passes + GoalsConceded, data=Football)
summary(LinearReg)
```

Base R functionality to plot regression results
```{r}
plot(LinearReg)
```

### Principal Component Analysis
With the *FactoMineR* package, perform PCA and plot the cumulative percentage of variance gained with each eigenvector
```{r}
library(FactoMineR)
FootballPC <- PCA(Football[,2:5])
summary(FootballPC)
#Plot cumulative variances as explained by the components
plot(1:4, FootballPC$eig$"cumulative percentage of variance",
     type = "b",
     xlab = "Axes 1,2,...,4",
     ylab = "Cum. % of variance")
FootballPC$eig$"cumulative percentage of variance"
```
