---
title: "R Cheatsheet"
output: html_document
---

Kate's personal log of interesting and useful R snippets

## Data Wrangling
### Data Input Basics
Creating a simple dataframe for example purposes
```{r}
Location <- c("New York", "London", "new york", "London")
Product <- c("Tea", "Coffee", "Coffee", NA)
Revenue <- c("USD 10000", "5000", "6,500", "$7850")
YearFounded <- c(1985, 2001, NA, NA)
SampleSet <- data.frame(Location, Product, Revenue, YearFounded, stringsAsFactors=FALSE)
SampleSet
```

Saving a copy of the dataframe as .csv (comma-separated) data
```{r}
write.csv(SampleSet, "SampleSet.csv")
```

Reading in .csv (comma-separated) data
```{r}
NewSampleSet <- read.csv("SampleSet.csv")
NewSampleSet
```

Reading in .tsv (tab-separated) data
```{r}
```

### Looking at Basic Structure
Show the dimensions of the dataframe
```{r}
dim(SampleSet)
```

See the distributions of values for a certain field
```{r}
table(SampleSet$Location)
```

List all unique values for a certain field
```{r}
unique(SampleSet$Location)
```

### Column Headings
Listing all column headings
```{r}
names(SampleSet)
```

Renaming a specific column heading
```{r}
```

Substituting new column names using rename()
```{r}
```

### Understanding structure of individual columns
See now many non-NA values are contained within a column
```{r}
length(which(!is.na(SampleSet$Product)))
```

### Cleaning up content
Duplicating a field in the dataset (but giving it a new name)
```{r}
SampleSet$OtherLocation <- SampleSet$Location
SampleSet
```

Delete a field from the dataset
```{r}
SampleSet$OtherLocation <- NULL
SampleSet
```

Convert all contents of a column to proper case
```{r}
SampleSet$Location

#Add a function to make proper case
properCase <- function(x) {
  lower <- tolower(x)
  s <- strsplit(lower, " ")[[1]]
  paste(toupper(substring(s, 1,1)), substring(s, 2),
      sep="", collapse=" ")
}
#Apply this function to a column
SampleSet$LocationProper <- sapply(SampleSet$Location, properCase)

SampleSet$LocationProper
```

Extract the digits in a column and make column into an integer column (by replacing all non-digit characters with blank space)
```{r}
SampleSet$Revenue <- as.integer(gsub("[^0-9]", "", SampleSet$Revenue))
SampleSet
```

## Plots
### Simple Plots in Base R
Show a histogram of the distribution of values
```{r}
hist(SampleSet$Revenue)
```

Same as previous, but with a lot more bins to the histogram
```{r}
hist(SampleSet$Revenue, breaks=100)
```

### Simple Plots Using ggplot2
